---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 3
#### Tutorial exercises 5.7

Create a dataframe with all the information.
```{r}
## create a dataframe
Age <- c(70,71,72,73,74,75)
Central_exposed_to_risk <- c(1000,1005,1010,1008,1006,998)
Observed_number_of_death <- c(80,90,95,105,115,125)
data_5.7 <- data.frame(Age,Central_exposed_to_risk,Observed_number_of_death)
data_5.7
```

Then we need to calculate the crude rate by $\mu_x=\frac{d_x}{E^{c}_{x}}$.
```{r}
# calculate crude rate qx
# and add this column to the original dataframe
data_5.7$Crude <- data_5.7$Observed_number_of_death/data_5.7$Central_exposed_to_risk
data_5.7
```

We need to obtain the graduated estimates of $\mu_x$ by using Gompertz' Law and maximum likelihood. Hence the graduated estimates is $\hat{\mu_x} = Bc^{x}$. 

We need to use maximum likelihood to get the value of parameters B and c. The model using in this question is a possion model, and the log-likelihood used for MLE is: $$lnL=\sum_{x=70}^{75}(-{E^{c}_x}Bc^x+d_xln({E^c_xBc^x})-ln(d_x!))$$

We then need to write the function so that we can use the "nlm" function in r to get the MLE of B and c.
```{r}
# rename variables
Ex_c <- data_5.7$Central_exposed_to_risk
dx <- data_5.7$Observed_number_of_death
x <- data_5.7$Age

# create a function
fn <- function(theta){
  # mu_x = Bc^x
  # B is theta[1], c is theta[2]
  log_likelihood = -sum(Ex_c*theta[1]*theta[2]^x)+sum(dx*log(Ex_c*theta[1]*theta[2]^x)) 
  return(-log_likelihood)
}
```

Before we use the "nlm" function, we need to have some initial values for B and c. Initial values are calculated by using linear regression. Since $\hat{\mu_x} = Bc^{x}$, then $ln\hat{\mu_x} = a + bx$. And $ln\hat{\mu_x}$ is approximately equal to $ln{q_x}$. Therefore, we can use $ln{q_x} = a + bx$ to get the initial values of B and c by calculating exp(a) and exp(b) respectively.

```{r}
## linelar regression to estimate starting point for B & c
qx <- data_5.7$Crude
log_qx <- log(qx)
lm(log_qx ~ x)
start_B <- exp(-8.67325)
start_c <- exp(0.08788)
```

Then sub the initial values of B and c into the "nlm" function to get the MLE of B and c.
```{r}
## use the starting point for MLE
MLE <- nlm(fn,c(start_B,start_c))
MLE
B <- MLE$estimate[1]
c <- MLE$estimate[2]
```

Now we can use $\hat{\mu_x} = Bc^{x}$ to calculate the graduated estimates of the force of mortality.
```{r}
## use the estimated B and c to calculate graduated estimate mu
data_5.7$Graduated <- B*c^x
data_5.7$Graduated
data_5.7
```




