---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 3
#### Ex 5.8

Create a dataframe with all the information.
```{r}
## create a dataframe
Age <- c(30:49)
Initial_exposed_to_risk <- c(70000,66672,68375,65420,61779,66091,68514,69560,
                             65000,66279,67300,65368,65391,62917,66537,62302,
                             62145,63856,61097,61110)
Observed_number_of_death <- c(39,43,34,31,23,50,48,43,48,47,62,63,84,86,120,
                              121,122,162,151,184)
data_5.8 <- data.frame(Age,Initial_exposed_to_risk,Observed_number_of_death)
data_5.8
```

Then we need to calculate the crude rate by $q_x=\frac{d_x}{E_x}$.
```{r}
## calcualte crude rate qx
# and add this column to the original dataframe
data_5.8$Crude <- Observed_number_of_death/Initial_exposed_to_risk
data_5.8
```

We need to obtain the graduated estimates of mortality rate using $ln\frac{\hat{q_x}}{1-\hat{q_x}}=a+bx$ and weighted least squares. Hence, the function we need to create to calculate the values of a and b is: $$f(a,b)=\sum_{x=30}^{49}{w_x[ln\frac{\hat{q_x}}{1-\hat{q_x}}-(a+bx)]^2}$$

For the equation above, $w_x$ is the weight, where in this case, $w_x=E_x$

we then need to write the function so that we can use "nlm" to get the estimates of a and b.

```{r}
## rename the variables
qx <- data_5.8$Crude
wx <- data_5.8$Initial_exposed_to_risk
x <- data_5.8$Age

## calculate y useing y = ln(qx/1-qx)
y <- log(qx/(1-qx))

## set up a function
fn <- function(theta){
  # a is theta[1], b is theta[2]
  f_ab <- sum(wx*(y-(theta[1]+theta[2]*x))^2)
  return(f_ab)
}
```

Before we use the "nlm" function, we need to have some initial values for a and b. Initial values are calculated by using linear regression. $ln\frac{\hat{q_x}}{1-\hat{q_x}}=a+bx$

```{r}
## linear regression 
lm(y~x)
start_a <- -10.8629
start_b <- 0.1003
```

Then sub the initial values of a and b into the "nlm" function to get the MLE of a and b.
```{r}
## Calculating estimated values of a and b
estimate <- nlm(fn,c(start_a,start_b))
estimate
a1 <- estimate$estimate[1]
b1 <- estimate$estimate[2]
```

Now we can now calculate the graduated estimates of mortality by ranging $ln\frac{\hat{q_x}}{1-\hat{q_x}}=a+bx$ into $\hat{q_x}=\frac{e^{(a+bx)}}{1+e^{(a+bx)}}$

```{r}
## use the estimated a,b to calculate graduated qx
## qx_grad = exp(a+bx)/(1+exp(a+bx))
data_5.8$Graduated_1 <- exp(a1+b1*x)/(1+exp(a1+b1*x))
data_5.8$Graduated_1
data_5.8
```

Try other inital values to see if we can get a better estimates of mortality.
```{r}
## try with different intital values
start_a = -10
start_b = 0

estimate <- nlm(fn,c(start_a,start_b))
estimate
a2 <- estimate$estimate[1]
b2 <- estimate$estimate[2]

## calculate new graduated value with new a and b
data_5.8$Graduated_2 <- exp(a2+b2*x)/(1+exp(a2+b2*x))
data_5.8$Graduated_2
data_5.8
```



