---
title: "Assignment 2"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1

##### (a) Estimate q30
First, we need to create a dataframe. Actual_Birthday_30 is the acutal date that each life has their 30th birthday, the policy anniversiry which is nearest each life's actual 30th birthday is PA_30, and the policy anniversiry that is nearest each life's actual 31st birthday is PA_31. 
<br/>
Date_of_Exit is determined based on whether each life dies, withdrawals or survives. For example, Life A survivals throughout PA_30 and PA_31, then its exit day is PA_31. Life G dies between PA_30 and PA_31, for initial exposed to risk, its exit day is still PA_31. For Life C, the calendar year of withdrawal is 1995, hence the exit day should be the policy anniversiry that has the same calendar year, which is PA_31.

```{r}
# create a dataframe
Life <- c('A','B','C','D','E','F','G','H','I','J')
Actual_Birthday_30 <- c('17/03/1994','06/05/1994','12/08/1994','27/10/1994',
                        '04/01/1995','18/04/1995','20/05/1995','04/07/1995',
                        '16/09/1995','11/12/1995')
PA_30 <- c('20/06/1994','06/08/1994','18/12/1994','04/01/1995','28/04/1995',
           '16/06/1995','29/10/1995','16/02/1995','22/08/1995','06/03/1996')
PA_31 <- c('20/06/1995','06/08/1995','18/12/1995','04/01/1996','28/04/1996',
           '16/06/1996','29/10/1996','16/02/1996','22/08/1996','06/03/1997')
Date_of_Death <- c('NA','12/06/1993','NA','NA','29/08/1996',"NA",'21/04/1996',
                   'NA','NA','17/02/1997')
Date_of_Withdrawal <- c('NA','NA','18/06/1995','NA','NA','12/12/1995','NA',
                        'NA','22/02/1997','NA')
## determine date of exit based on date of death, date of withdrawal
Date_of_Exit <- c('20/06/1995','06/08/1994','18/12/1995','04/01/1996','28/04/1996',
                  '16/06/1995','29/10/1996','16/02/1996','22/08/1996','06/03/1997')
data <- data.frame(Life,Actual_Birthday_30,PA_30,PA_31,Date_of_Death,
                   Date_of_Withdrawal,Date_of_Exit)
data
```

Then we can calculate initial exposed to risk which can be calculated as the difference between Date_of_Exit and PA_30.

```{r}
## calculate initial exposed to risk
data$Initial_Exposed_to_Risk <- as.Date(as.character(Date_of_Exit),format = "%d/%m/%Y")-
                                as.Date(as.character(PA_30),format = "%d/%m/%Y")
data
```

The total initial exposed to risk for 10 lives is the sum of each initial exposed to risk.
```{r}
## Ex
E30 <- sum(data$Initial_Exposed_to_Risk > 0)
```

To calculate totla number of death, it can be seen that there are four deaths. However, only Life G and Life J's deaths lie between PA_30 and PA_31. Hence, $d_x$ is 2.
```{r}
## count number of death
d30 <- 2
```

Hence, estimated $q_x$ can be calculated as $q_{30}=\frac{d_{30}}{E_{30}}$.
```{r}
## calculate q30
q30 <- d30/E30
q30
```

Therefore, $\hat{q_{30}}$ is 0.25.

##### (b) MLE of q30
First, we need to calcualte the central exposed to risk for each Life. The only difference from initial exposed to risk is that for death, the exposure is from PA_30 to the date of death. For life that withdrawals, the exposure is the same. Hence, the new exit date only changes for those who has death between PA_30 and PA_31.
```{r}
data$Date_of_Exit_new <- c('20/06/1995','06/08/1994','18/12/1995','04/01/1996',
                      '28/04/1996','16/06/1995','21/04/1996','16/02/1996',
                      '22/08/1996','17/02/1997')
data$Central_Exposed_to_Risk <- as.Date(as.character(data$Date_of_Exit_new),format = "%d/%m/%Y")-
                                as.Date(as.character(PA_30),format = "%d/%m/%Y")

data
```

Now, we write down the likelihood function for $q_30$ based on the information. We have 3 lives that survive throughout the period, 2 lives die between PA_30 and PA_31, 1 life dies before PA_30, and 1 life dies after PA_31, 2 lives have withdrawal age at 31 and one life have withdrawal age of 30. Hence, the likelihood function is $$L=(p_{30})^6{_\frac{175}{365.25}q_{30}}{_\frac{348}{365.25}q_{30}}$$ 

```{r}
fn <- function(qx){
  Log <- (1-qx)^6*(1-(1-qx)^(175/365.25))*(1-(1-qx)^(348/365.25))
  return(-Log)
}
nlm(fn,p=0.4)
```
The MLE of $q_{30}$ is 0.2584841.





