---
title: "Assignment 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 3
#### Ex 5.9

Create a dataframe with all the information.
```{r}
## create a timeframe
Age <- c(47:67)
Initial_exposed_to_risk <- c(166,187,218,243,276,302,347,390,430,494,558,
                             628,701,813,917,1040,1182,1299,1432,1596,1752)
Observed_number_of_death <- c(2,2,4,6,2,4,7,3,9,9,8,11,14,18,18,24,30,43,41,
                              54,64)
Standard_table_figures <- c(0.00505,0.00570,0.00644,0.00728,0.00826,0.00930,
                            0.01051,0.01184,0.01331,0.01492,0.01668,0.01859,
                            0.02065,0.02287,0.02525,0.02778,0.03049,0.03339,
                            0.03648,0.03978,0.04332)
data_5.9 <- data.frame(Age,Initial_exposed_to_risk,Observed_number_of_death,
                       Standard_table_figures)
data_5.9
```

Then we need to calculate the crude rate by $q_x=\frac{d_x}{E_x}$.
```{r}
## calcuate crude rate qx
data_5.9$Crude <- Observed_number_of_death/Initial_exposed_to_risk
data_5.9
```

we need to obtain the graduated estimates of mortality rate by using standard table figures with $\hat{q_x}=a+b\hat{q_x}$ and weighted least squares. Hence, the function we need to create is $$f(a,b)=\sum_{x=47}^{67}{w_x[\hat{q_x}-(a+bq^s_x)]^2}$$

For the equation above, $w_x$ is the weight, where in this case, $w_x=E_x$

we then need to write the function so that we can use "nlm" to get the estimates of a and b.

```{r}
## rename variables
qx <- data_5.9$Crude
qx_s <- data_5.9$Standard_table_figures
Wx <- data_5.9$Initial_exposed_to_risk

## create a function
fn <- function(theta){
  # qx = a + b*qx_s
  # a is theta[1], b is theta[2]
  f_ab <- sum(Wx*(qx-(theta[1]+theta[2]*qx_s))^2)
  return(f_ab)
}
```

Before we use the "nlm" function, we need to have some initial values for a and b. Initial values are calculated by using linear regression. $\hat{q_x}=a+b\hat{q_x}$

```{r}
## linear regression to get the starting value for estimate a,b
lm(qx~qx_s)
start_a <- 0.009279
start_b <- 0.570217
```

Then sub the initial values of a and b into the "nlm" function to get the estimated values of a and b.
```{r}
## Calculating estimated values of a and b
estimate <- nlm(fn,c(start_a,start_b))
estimate
a <- estimate$estimate[1]
b <- estimate$estimate[2]
```

Now we can calculating the graduated estimates of mortality by $\hat{q_x}=a+b\hat{q_x}$
```{r}
## calculate the graduated rate
data_5.9$Graduated <- a + b*qx_s
data_5.9$Graduated
data_5.9
```






